---
title: "Lab2"
author: "Kate McGeath"
date: "9/2/2022"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Tasks

## Task 1

```{r}
getwd()
```

## Task 2

```{r}
epagas = read.csv("EPAGAS.csv")
head(epagas)
```

## Task 3

Make the object mgp: the number of miles per gallon vector
```{r}
mpg = epagas$MPG
head(mpg)
```

Transform mpg to z and verify results:
```{r}
z = (mpg - mean(mpg)) / sd(mpg)
head(z)
print(paste0("mean = ", round(mean(z), digits = 4)), quote = FALSE)
print(paste0("variance = ", var(z)), quote = FALSE)
print(paste0("standard deviation = ", sd(z)), quote = FALSE)
```

Values of mpg that are possible outliers:
```{r}
mpg[abs(z) >= 2 & abs(z) <= 3]
```

Values of mpg defined as outliers:
```{r}
mpg[abs(z) > 3]
```

```{r}
mpgcol = ifelse(abs(z) > 3, "red",
                ifelse(abs(z) >= 2 & abs(z) <= 3, "blue", "black"))

library(lattice)
dotplot(mpg, col = mpgcol, main = "Plot of mpg values")
```

## Task 4

```{r}
boxplot(mpg, main = "MPG Boxplot", xlab = "MPG", col = "black", horizontal = TRUE, notch = TRUE)
```

Chebyshev's Theorm says at least

1 - (1 / k^2)

of the data will fall within k standard deviations of the mean.
With k = 2,
```{r}
k = 2
cheb = 1 - 1 / (k^2)
cheb
```

The Proportion/Percentage of data within 2 standard deviations of the mean:
```{r}
tempData<-epagas[abs(z)<2,"MPG"]
length(tempData)/length(epagas$MPG)*100
```

Chebyshev would agree because more than 3/4ths of the data lies within 2 standard deviations.

According to the Empirical Rule, 95% will fall within 2 standard deviations of the mean.
Therefore, the Empirical rule is valid.
